Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-01-12 main.tex

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 \textbf{CHECK AGAIN} for reported cases and deaths from COVID-19. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 CHECK AGAIN forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. In addition, we filtered out erroneous forecasts characterised by a median forecast of 0 and an actual observed value of larger than 150. All predictive quantiles were truncated at 0. We applied the log-transformation after adding 1 to all predictions and observed values. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
    \caption{Correlations for median forecasts, scores, and rankings on the natural and logarithmic scale. A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values. B: Average Spearman rank correlation of scores for individual forecasts from all models. For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model. Then, the rank correlation was computed between the scores for all models on the natural scale vs. on the log scale. All rank correlations were averaged across locations, and target types and stratified by horizon and target type. C: Correlation between relative skill scores. For every forecast horizon and target type, a separate rel. skill score was computed per model using pairwise comparisons. The plot shows the correlation between the rel. skill scores on the natural vs. on the log scale.}
    \label{fig:HUB-cors}
\end{figure}

@@

\begin{table}
\centering
\begin{tabular}{ccrrrrrr}
\toprule
Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
\midrule
all & all & -1.093 & 0.963 & 0.390 & 0.001 & -0.995 & 0.209\\
\addlinespace
all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.047 & 0.201\\
all & Deaths & -0.885 & 0.868 & 0.430 & -0.021 & -7.564 & 0.826\\
\addlinespace
1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.318 & 0.088\\
2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.111 & 0.164\\
3 & all & -1.001 & 0.984 & 0.408 & 0.008 & -1.283 & 0.255\\
4 & all & -0.757 & 0.986 & 0.467 & 0.016 & -1.041 & 0.312\\
\addlinespace
1 & Cases & -0.861 & 0.876 & 0.433 & -0.024 & 0.785 & 0.087\\
2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.115 & 0.238\\
4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.665 & 0.296\\
\addlinespace
1 & Deaths & -1.145 & 0.831 & 0.376 & -0.035 & 0.458 & 0.048\\
2 & Deaths & -0.980 & 0.866 & 0.415 & -0.027 & 0.445 & 0.084\\
3 & Deaths & -0.809 & 0.885 & 0.441 & -0.016 & -8.262 & 0.903\\
4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
\bottomrule
\end{tabular}
\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$. The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
}
\label{tab:HUB-regression}
\end{table}

@@

To check this, we ran the regressions
\begin{equation*}
    \text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
\end{equation*}
where $F_{\log}$ is the predictive distribution for $\log(y)$, and
\begin{equation*}
    \text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
\end{equation*} 
where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale. As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$). The results of the three regressions are also displayed in Figure \ref{fig:HUB-cors}. It appears that the log transformation is most suitable to stabilise the WIS values in this empirical example.

@@

For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}B). When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably. The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}C) and possibly tends to increase with increasing forecast horizon.

When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D). We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases. 
Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F). On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.

@@

\begin{table}[h!]
    \centering

\begin{tabular}{lllcc}
\toprule
target\_type & quantity & measure & natural & log\\
\midrule
Cases & Observations & mean & 62004 & 9.19\\
Cases & Observations & sd & 171723 & 2.10\\
Cases & Observations & var & 29488651564 & 4.42\\
\addlinespace
Deaths & Observations & mean & 223 & 3.90\\
Deaths & Observations & sd & 441 & 1.96\\
Deaths & Observations & var & 194221 & 3.85\\
\addlinespace
\hline
\addlinespace
Cases & WIS & mean & 15840 & 0.27\\
Cases & WIS & sd & 53117 & 0.28\\
\addlinespace
Deaths & WIS & mean & 31 & 0.23\\
Deaths & WIS & sd & 65 & 0.28\\
\bottomrule
\end{tabular}
    \caption{Summary statistics for observations and scores for forecasts from the ECDC data set.}
    \label{tab:HUB-summary}
\end{table}

>>>>

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 for reported cases and deaths from COVID-19. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. In addition, we filtered out erroneous forecasts characterised by a median forecast of 0 and an actual observed value of larger than 150. All predictive quantiles were truncated at 0. We applied the log-transformation after adding 1 to all predictions and observed values. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
    \caption{Relationship between median forecasts and scores. Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble. Shown in red are the regression lines discussed in \ref{sec:regression} and \ref{tab:HUB-regression}. A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values. B: Same as A, with scores obtained after applying a log-transformation to the data. C: Same as A, with scores obtained after applying a square-root-transformation to the data.} 
    \label{fig:HUB-regression}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
    \caption{Correlations of rankings on the natural and logarithmic scale. A: Average Spearman rank correlation of scores for individual forecasts from all models. For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model. Then, the rank correlation was computed between the scores for all models on the natural scale vs. on the log scale. All rank correlations were averaged across locations, and target types and stratified by horizon and target type. B: Correlation between relative skill scores. For every forecast horizon and target type, a separate rel. skill score was computed per model using pairwise comparisons. The plot shows the correlation between the rel. skill scores on the natural vs. on the log scale.}
    \label{fig:HUB-cors}
\end{figure}

@@

\begin{table}
\centering
\begin{tabular}{ccrrrrrr}
\toprule
Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
\midrule
all & all & -1.093 & 0.963 & 0.390 & 0.001 & -0.995 & 0.209\\
\addlinespace
all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.047 & 0.201\\
all & Deaths & -0.885 & 0.868 & 0.430 & -0.021 & -7.564 & 0.826\\
\addlinespace
1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.318 & 0.088\\
2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.111 & 0.164\\
3 & all & -1.001 & 0.984 & 0.408 & 0.008 & -1.283 & 0.255\\
4 & all & -0.757 & 0.986 & 0.467 & 0.016 & -1.041 & 0.312\\
\addlinespace
1 & Cases & -0.861 & 0.876 & 0.433 & -0.024 & 0.785 & 0.087\\
2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.115 & 0.238\\
4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.665 & 0.296\\
\addlinespace
1 & Deaths & -1.145 & 0.831 & 0.376 & -0.035 & 0.458 & 0.048\\
2 & Deaths & -0.980 & 0.866 & 0.415 & -0.027 & 0.445 & 0.084\\
3 & Deaths & -0.809 & 0.885 & 0.441 & -0.016 & -8.262 & 0.903\\
4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
\bottomrule
\bottomrule
\end{tabular}
\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$. The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
}
\label{tab:HUB-regression}
\end{table}

@@

To check this, we ran the regressions
\begin{equation*}
    \text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
\end{equation*}
where $F_{\log}$ is the predictive distribution for $\log(y)$, and
\begin{equation*}
    \text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
\end{equation*} 
where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale. As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$). The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}. It appears that the log transformation is most suitable to stabilise the WIS values in this empirical example.

@@

For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A). When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably. The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}B) and possibly tends to increase with increasing forecast horizon.

When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D). We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases. 
Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F). On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.

@@

\begin{figure}[h!]
\centering
\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors-sqrt.png}
\caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to \textit{sqrt}-transformed predictions and observations. We denote the predicted value by $\hat{y}$ and display errors as a function of the ratio of observed and predicted value. A: x-axis shown on a linear scale. B: x-axis shown on a logarithmic scale.}
\label{fig:SAPE-sqrt}
\end{figure}

\begin{table}[h!]
    \centering
\begin{tabular}{lllcc}
\toprule
target\_type & quantity & measure & natural & log\\
\midrule
Cases & Observations & mean & 62004 & 9.19\\
Cases & Observations & sd & 171723 & 2.10\\
Cases & Observations & var & 29488651564 & 4.42\\
\addlinespace
Deaths & Observations & mean & 223 & 3.90\\
Deaths & Observations & sd & 441 & 1.96\\
Deaths & Observations & var & 194221 & 3.85\\
\addlinespace
\hline
\addlinespace
Cases & WIS & mean & 15840 & 0.27\\
Cases & WIS & sd & 53117 & 0.28\\
\addlinespace
Deaths & WIS & mean & 31 & 0.23\\
Deaths & WIS & sd & 65 & 0.28\\
\bottomrule
\end{tabular}
    \caption{Summary statistics for observations and scores for forecasts from the ECDC data set.}
    \label{tab:HUB-summary}
\end{table}
